---
title: "10/28 Notes"
author: "Sepehr Akbari"
date: "10/28/2024"
format: html
embed-resources: true
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
```

## Two-sample t-test example

Suppose we have a random sample of 100 diamonds.

```{r}
set.seed(0)

diamonds_sample <- diamonds %>% 
  slice_sample(n = 100)
```

Are colors D and E more expensive than the others? 

First, we need a column that is gonna indicate whether each diamond is D/E or not. 

```{r}
diamonds_sample <- diamonds_sample %>% 
  mutate(
    clear = color %in% c("D", "E")
  )

glimpse(diamond_sample)
```

Now we can do a two-sample t-test.

Null: The prices are the same for T and F

Alternative: the prices are different

```{r}
t.test(diamonds_sample$price ~ diamonds_sample$clear)
```

The p-value is small, not infinitesimally small, but small enough to reject the null hypothesis. The prices in the two groups in the population appear different.

Can we conclude the clear ones are more expensive? No it appears that the clear ones tend to be less expensive in the population (interesting).

## Testing for a categorical variable

Let's get a new sample of diamonds. This time let's take 250.

```{r}
diamonds_sample2 <- diamonds %>% 
  slice_sample(n = 250)
```

A store with this stock claims that in the larger population, the cuts of diamonds have proportionsas follows:

```{r}
props <- c(0.05, 0.1, 0.2, 0.25, 0.4)
cuts <- c("Fair", "Good", "Very Good", "Premium", "Ideal")

df <- data.frame(cuts, props)
gt::gt(df)
```

Let's find the proportions in the sample.

```{r}
table(diamonds_sample2$cut)
table(diamonds_sample2$cut) / 250
```

Is the store's claim plausible based on this data?

First, notice that we can't really do a confidence interval for a catagerical variable. We can only do a sig-test.

Null: the props claimed by the store are correct for the population.

Alternative: the props are different.

We'll test this using a **chi-squared goodness of fit test**, which compares the observed counts with the ones expected if the null is true.

```{r}
expected <- props * 250
table(diamonds_sample2$cut)

chisq.test(table(diamonds_sample2$cut), p = props)
```

The Chi-squared statistic measures how far observed counts are from expected ones across multiple categories.

$$ \chi^2 = \sum\frac{(O - E)^2}{E} $$
Where O is observed and E is expected. In R, we run the with `chisq.test()`, which needs a vector of counts and another vector of expected proportions.

```{r}
chisq.test(table(diamonds_sample2$cut), p = props) # p = x is necessary
```

Where p is a vector of expected proportions. Here p is very large. The null hypothesis is consistent with the data. We draw no further conclusions.



