---
title: "11/15 Notes"
author: "Sepehr Akbari"
date: "11/15/2024"
format: html
embed-resources: true
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(readxl)
```


## Finding files nested inside folders

My parenting dataset lies inside a folder called `data`. which is indside my project.

A bad idea is to use the `setwd()` command which forces a change to the working directory.

```{r}
library(here)
```

We build file paths using the `here()` command in the `here` package. For me `parenting.xlsx` sits inside my `Data` in the project folder

```{r}
path <- here("Data", "parenting.xlsx")

parenting <- read_excel(path)
```

## A bit more prep for spotifyr project

Its not a good idea to download artist data inside your quarto for the final project. Please don't. Instead, download it once and then write in to a csv file with the following code `write_csv`. Like this:

```{r}
#| eval: false

write_csv(amy, "amy.csv")
```

now you can start your project #2 quarto with a `read_csv` command to import the set.

## Intro to simple linear regression

The `parenting` set includes 100 observations of a parent's mood after a night in which they and the baby slept a certain amount. BTW, this set is credited to Danielle Navarro, a big name in R and DSCI.

Is her mood related to the amount of sleep the baby got? Note that this is a question about two quantitatiive variables.

```{r}
ggplot(
  parenting,
  aes(x = baby.sleep,
      y = dan.grump)
) +
  geom_point()
```

The trend is linear, lets add a line of best fit.

```{r}
ggplot(
  parenting,
  aes(x = baby.sleep,
      y = dan.grump)
) +
  geom_point() +
  geom_smooth(method = "lm",
              se = FALSE)
```

Let's get the equation of this line.

```{r}
lm(dan.grump ~ baby.sleep, data = parenting)
```

The model says `dan.grump` = 85.8 - 2.7 * `baby.sleep` + (random noise).

