---
title: "10/30 Notes"
author: "Sepehr Akbari"
date: "10/30/2024"
format: html
embed-resources: true
editor_options: 
  chunk_output_type: console
---

# 10/30 Notes

```{r}
library(tidyverse)
library(modeldata)
```

Let's consider another sample of diamonds.

```{r}
set.seed(1234)

diamonds_sample <- diamonds %>% 
  slice_sample(n = 250)
```

In the larger population which we ordinarily don't have, are the different colors equally likely? Use the sample data to decide.

Null: the proportions in the population are all equal.

Alternative: the proportions in the population are not all equal.

Let's start with a visualization.

```{r}
ggplot(
  diamonds_sample,
  aes(x = color)
) +
  geom_bar()
```

It looks like the colors aren't evenly distributed but this could be due to random chance in the sampling procedure. A p-value will tell us how likely that is.

Next, let's get observed counts.

```{r}
obs <- table(diamonds_sample$color)
```

Now let's run our test on these observed counts.

```{r}
chisq.test(obs, 
           p = rep(1/7, 7)) # repeats 1/7, 7 times
```

if we accept the null, we draw no further conclusion, but if we reject the null, we conclude that the colors are not evenly distributed in the larger population. The infentestimal p-value indicates small evidence against the null hypothesis. We conclude that diamond colors are not evenly distributed in the larger population.

A little shortcut:

```{r}
chisq.test(obs) # same as above
```

the default null is equal proportions in the population. But its better to be explicit about the null hypothesis.

## Two-variable inference: categorical data

We're now considering whether two variables of interest are independent or not. That is, does knowing the value of one of the variables give us information about the value of the other variable?

We'll start with two categorical variables, using the **chi-squared test of independence**.

Example: the `stackoverflow` set.

```{r}
glimpse(stackoverflow)
```

We view this set as a random sample from a much larger population. Note this is an assumption!

Question: are the variables `Remote` and `Country` independent? That is if I know your country, does that help me guess whether you work remotely?

null: the variables are independent

alternative: the variables are not independent

```{r}
table(stackoverflow$Remote, 
      stackoverflow$Country) # a numerical summary
```

To visualize two categorical variables, we can use a bar chart.

```{r}
ggplot(
  stackoverflow,
  aes(x = Country, fill = Remote)
) +
  geom_bar(position = "fill") +
  scale_fill_brewer(palette = "Dark2")
```

# 11/1 Notes

We were looking at whether `Remote` and `Country` are independent.

To run a chi-squared test of independence, we need to get a table of observed counts.

```{r}
s_table <- table(stackoverflow$Remote, 
                 stackoverflow$Country)
```

To test the hypotheses, compare observed counts in this table to the counts we'd expect if the null were true.

```{r}
chisq.test(s_table) # run the test for independence
```

The very low p-value ($3.3\times10^{-16}$) provides very strong evidence against the null hypothesis. We conclude that `Remote` and `Country` are not independent. Meaning `Country` gives information about `Remote`.

Notice that the differences while statistically significant aren't large in magnitude. The **chi-squared test for independence doesn't measure size of difference, and it doesn't say anything about which groups might be different**

## Another example: `scat`

Do different species poop in different locations?

```{r}
glimpse(scat)
```

Null: the variables are independent. Knowing location doesn't give information about species and vice versa.

Alternative: the variables are not independent. Knowing location gives information about species and vice versa.

Let's get a visualization.

```{r}
ggplot(
  scat,
  aes(x = Location, fill = Species)
) +
  geom_bar(position = "fill") +
  scale_fill_brewer(palette = "Dark2") +
  theme_minimal()
```

Now for a table.

```{r}
poop_table <- table(scat$Species,
                    scat$Location)

poop_table
```

And the test itself:

```{r}
chisq.test(poop_table)
```

The p-value is only kinda small. Typically a cutoff of 0.05 is used, but this isn't gospel. Always view the p-value as giving the strength of evidence against the null hypothesis. We fail to reject the null.

However, in this case, we still wonder if the variables are dependent in the population. It's very possible we just didn't detect the differences, because of the small sample.

*Next-up: **ANOVA**. Which tests whether a quantitative variable is independent of a categorical variable.*
