---
title: "11/18 Notes"
author: "Sepehr Akbari"
date: "11/18/2024"
format: html
embed-resources: true
editor_options: 
  chunk_output_type: console
---

## More parenting regrssion

```{r}
library(tidyverse)
library(readxl)

parenting <- read_excel("Data/parenting.xlsx")
```

We view this as a random sample, imagining a much larger population of days with the same variables. We might be interested in whether there is a relationship between two quantitative variables, like `baby.sleep` and `dan.grump`. Linear regression does this, specifically testing for a linear relationship.

NULL: The average of `dan.grump` (the response variable) is not linearly related to the value of `baby.sleep` (the explanatory variable).

ALT: The average of `dan.grump` is linearly related to the value of `baby.sleep`. Specifically, the slope of the regression line is not zero.

As usual you should set hypotheses before looking at the sample data. You should also make a plot (we did this on 11/15).

```{r}
lm(dan.grump ~ baby.sleep, data = parenting) %>% 
  summary()
```

