---
title: "12/2 Notes"
author: "Sepehr Akbari"
date: "12/2/2024"
format: html
embed-resources: true
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(readxl)
library(here)

path <- here("data", "parenting.xlsx")

parenting <- read_excel(path)
```

## Parenting model comparison

```{r}
model_baby <- lm(dan.grump ~ baby.sleep, data = parenting)
model_dan <- lm(dan.grump ~ dan.sleep, data = parenting)
model_full <- lm(dan.grump ~ baby.sleep + dan.sleep, data = parenting)
```

A generally poor way of comparing is using the $R^2$. First, this measures the spread of the data about the regression line, not whether the model is well-fit. Second, adding variables to a regression model (like how we added `dan.sleep` to `model_baby`) can only increase the $R^2$ value, even if the new variable is just noise.

An improvement is **adjusted $R^2$**. This penalizes the $R^2$ value for complex models. There are others, for instance AIC.

```{r}
summary(model_baby)
summary(model_dan)
summary(model_full)
```

We prefer `model_dan` over the others.

Warning: if you are using $R^2$ and/or p-values for model selection, don't also do a hypothesis test.

## Categorical Variables

Let's look at `iris` and restrict ourselves to virginica and versicolors.

```{r}
iris_sm <- iris %>% 
  filter(Species != "setosa")

ggplot(
  iris_sm,
  aes(x = Sepal.Width, 
      y = Sepal.Length,
      color = Species)
) +
  geom_point() +
  geom_smooth(method = "lm",
              se = FALSE) +
  scale_color_brewer(palette = "Dark2")
```

A parallel-slopes model makes sense. There is a linear trend in each group with vary similar-looking slopes.

```{r}
model_par <- lm(Sepal.Length ~ Sepal.Width + Species, data = iris_sm)
```










