---
author: "Sepehr Akbari"
date: "20 march 2024"
format: pdf
---

###### Sepehr Akbari

###### March 22, 2024

# The central limit theorem

## Problem 1

A certain population has mean $\mu = 27.3$ and standard deviation $\sigma = 4.2$. For a sample of size $n = 36$,

(a) Determine the mean, variance, and standard deviation of the sample mean $x$.

**Answer:**

```{r}
mu <- 27.3
sigma <- 4.2
n <- 36

cat("mean of x:",mu,"\n")
cat("variance of x:",(sigma ^ 2 / n),"\n")
cat("standard deviation of x:",sqrt((sigma ^ 2 / n)),"\n")
```

(b) Determine the probability that $x ≤ 26.0$.

**Answer:**

```{r}
mu <- 27.3
sigma <- 4.2
n <- 36
stndDev <- (sigma / sqrt(n))

z <- (26.0 - mu) / stndDev

pnorm(z)
```

(c) Determine the probability that $x ≥ 25.9$.

**Answer:**

```{r}
mu <- 27.3
sigma <- 4.2
n <- 36
stndDev <- (sigma / sqrt(n))

z <- (25.9 - mu) / stndDev

1 - pnorm(z)
```

## Problem 2

At a local grocery, apples have mean weight $0.620$ pounds with standard deviation $0.165$ pounds. The distribution is approximately normal.

(a) What is the probability that a randomly-selected apple weighs more than $0.650$ pounds?

**Answer:**

```{r}
mu <- 0.62
sigma <- 0.165

cat("The probablity is:",1 - pnorm(0.65,mu,sigma))
```

(b) What is the probability that $10$ randomly-selected apples weigh more than $0.650$ pounds, on average?

**Answer:**

```{r}
mu <- 0.62
sigma <- 0.165
n = 10
stndDev <- (sigma / sqrt(n))

z <- (0.65 - mu) / stndDev

cat("The probablity is:",1 - pnorm(z))
```

(c) What is the probability that $50$ randomly-selected apples weigh more than $0.650$ pounds, on average?

**Answer:**

```{r}
mu <- 0.62
sigma <- 0.165
n = 50
stndDev <- (sigma / sqrt(n))

z <- (0.65 - mu) / stndDev

cat("The probablity is:",1 - pnorm(z))
```

(d) What is the probability that $500$ randomly-selected apples weigh more than $0.650$ pounds, on average?

**Answer:**

```{r}
mu <- 0.62
sigma <- 0.165
n = 500
stndDev <- (sigma / sqrt(n))

z <- (0.65 - mu) / stndDev

pnorm(z)

cat("The probablity is about:",1 - 0.999)
```

## Problem 3

Lengths of eruptions of the Old Faithful geyser are approximately normally distributed with mean $3.49$ minutes and standard deviation $1.14$ minutes.

(a) Which is more likely, a single eruption longer than $3.20$ minutes or $20$ eruptions with mean greater than $3.20$ minutes? Compute both probabilities.

**Answer:**

```{r}
mu <- 3.49
sigma <- 1.14
n = 20
x = 3.20
stndDev <- (sigma / sqrt(n))

single_eruption <- 1 - pnorm(x,mu,sigma)
multi_eruption <- 1 - pnorm((x - mu) / stndDev)

print(single_eruption)
print(multi_eruption)

if(single_eruption > multi_eruption){
  cat("The probability of a single eruption of 3.2 minutes or greater is higher, as:",single_eruption,">",multi_eruption,".")
}else if (multi_eruption > single_eruption){
  cat("The probability of the mean of 20 eruptions being greater than 3.2 minutes is higher, as:",multi_eruption,">",single_eruption,".")
}else{
  cat("They are equally probable.")
}
```

(b) Which is more likely, a single eruption longer than $3.60$ minutes or $20$ eruptions with mean greater than $3.60$ minutes? Justify your answer without computing probabilities.

**Answer:**

```{r}
mu <- 3.49
sigma <- 1.14
n = 20
x = 3.60
stndDev <- (sigma / sqrt(n))

single_eruption <- 1 - pnorm(x,mu,sigma)
multi_eruption <- 1 - pnorm((x - mu) / stndDev)

print(single_eruption)
print(multi_eruption)

if(single_eruption > multi_eruption){
  cat("The probability of a single eruption of 3.6 minutes or greater is higher, as:",single_eruption,">",multi_eruption,".")
}else if (multi_eruption > single_eruption){
  cat("The probability of the mean of 20 eruptions being greater than 3.6 minutes is higher, as:",multi_eruption,">",single_eruption,".")
}else{
  cat("They are equally probable.")
}
```
