---
author: "Sepehr Akbari"
date: "04 march 2024"
format: pdf
---

# Uniform Random Variables

## Problem 1

Compute the following in N(0,1).

(a) $P(Z < -0.6)$

```{r}
pnorm(-0.6)
```

(b) $P(Z > 1.3)$

```{r}
1 - pnorm(1.3)
```

(c) $P(−1.2≤ Z ≤ 2.1)$

```{r}
pnorm(2.1) - pnorm(-1.2)
```

## Problem 2

The speed of a car on cruise control has a normal distribution with mean $\mu = 72$ mph and standard deviation $σ = 1.1$ mph.

(a) Find the Z-score corresponding to a speed of $70$ mph.

Answer:

$$
\frac{70-72}{1.1} = -1.82
$$

(b) Compute the probability that the car is traveling more than $70$ mph at a random moment using the Z-score from part (a).

```{r}
pnorm(-1.82)
```

(c) Check your answer from part (b) using $N(72,1.1^2)$.

```{r}
pnorm(70,72,1.1)
```

{{< pagebreak >}}

# Inverse normal calculations

## Problem 1

Find the z-score that has $44\%$ of the distribution to its left.

Answer:

```{r}
qnorm(0.44)
```

## Problem 2

Find the value in $N(12,32)$ that has 87% of the distribution to its right.

Answer:

```{r}
qnorm(1 - 0.87,12,3)
```

## Problem 3

The average daily high temperature in June in LA is $77$F with a standard deviation of $5$F. Suppose that the temperatures in June closely follow a normal distribution.

(a) What is the probability of observing an $83$F temperature or higher in LA during a randomly chosen day in June?

Answer:

```{r}
1 - pnorm(83,77,5)
```

(b) How cool are the coldest $10\%$ of the days (days with lowest high temperature) during June in LA?

Answer:

```{r}
qnorm(0.1,77,5)
```

## Problem 4

Small bags of chips have weights that are normally distributed with mean $μ = 1.55$ oz and standard deviation $σ = 0.06$ oz.

(a) What is the probability that a randomly-selected bag of chips weighs less than 1.50 oz?

Answer:

```{r}
pnorm(1.5, 1.55, 0.06)
```

(b) What is the 98th percentile of weights?

Answer:

```{r}
qnorm(0.98, 1.55, 0.06)
```
