---
author: "Sepehr Akbari"
date: "12 April 2024"
format: pdf
---

###### Sepehr Akbari

###### April 12th, 2024

# Hypothesis testing with t

## Instructions

For each of the following problems,

(a) Write null and alternative hypotheses appropriate to this study.

(b) Compute the t-score of the sample mean.

(c) Compute the p-value of the sample mean.

(d) Are the results statistically significant at level $α = 0.05$?

(e) What conclusions, if any, can be drawn from this study? Answer in ordinary human language.

(f) Verify your calculations using the t.test( ) function.

## Problem 1

A fluorescent lamp manufacturer advertises that the mean life of their lamps is $10,000$ hours. You worry that it’s less. Use the lamp data set, available on Moodle, to test this claim at significance level $α = 0.05$.

**Answer:**

Loading the dataset:

```{r}
library(readxl)
lamp <- read_excel("/Users/sepehrakbari/Documents/LFC/Semester 2/MATH 150/DSs/lamps.xlsx")
```

*(a) Null & Alternative Hypotheses*

Null Hypotheses: $H_0 = 10,000$

Alternative Hypotheses: $H_1 < 10,000$

*(b) t-score*

```{r}
populationMean <- 10000
mean <- mean(lamp$hours_of_use)
stndDev <- sd(lamp$hours_of_use)
count <- length(lamp$hours_of_use)

t_score <- (mean - populationMean) / (stndDev / sqrt(count))

cat("t-score:", t_score, "\n")
```

*(c) p-value*

```{r}
p_value <- pt(t_score, count - 1)

cat("p-value:", p_value, "\n")
```

*(d) statistical significancy test*

```{r}
alpha <- 0.05 

if (p_value < alpha) {
  cat("Statistically significant (H1).")
} else {
  cat("Not statistically significant (H0).")
}
```

*(e) conclusion*

We cannot confirm with this data that the average life of a lamp is less than 10,000 hours. Either the null hypotheses ($H_0 = 10000$) is true, or a larger sample size or a lower significance level might be needed for a clearer picture.

*(f) verification*

```{r}
t.test(lamp$hours_of_use, mu = 10000, alternative = "less")
```

## Problem 2

A guidebook says that the average time between eruptions of the Wyoming’s Old Faithful geyser is $75$ minutes. Use built-in R data set faithful to test this clasim at significance level $α = 0.05$.

**Answer:**

*(a) Null & Alternative Hypotheses*

Null Hypotheses: $H_0 = 75$

Alternative Hypotheses: $H_1 \neq 75$

*(b) t-score*

```{r}
populationMean <- 75
mean <- mean(faithful$waiting)
stndDev <- sd(faithful$waiting)
count <- length(faithful$waiting)

t_score <- (mean - populationMean) / (stndDev / sqrt(count))

cat("t-score:", t_score, "\n")
```

*(c) p-value*

```{r}
p_value <- 2 * pt(t_score, count - 1)

cat("p-value:", p_value, "\n")
```

*(d) statistical significancy test*

```{r}
alpha <- 0.05 

if (p_value < alpha) {
  cat("Statistically significant (H1).")
} else {
  cat("Not statistically significant (H0).")
}
```

*(e) conclusion*

The calculation shows that there is enough evidence supporting that the mean eruption in the Old Faithful geyser is indeed $75$ minutes.

*(f) verification*

```{r}
t.test(faithful$waiting, mu = 75, alternative = "two.sided")
```