---
author: "Sepehr Akbari"
date: "05 April 2024"
format: pdf
---

###### Sepehr Akbari

###### April 5th, 2024

# Introduction to hypothesis testing

## Introduction:

For each of the following problems,

(a) Write null and alternative hypotheses appropriate to this study.

(b) Compute the z-score of the sample mean.

(c) Compute the p-value of the sample mean.

(d) Are the results statistically significant at level $α = 0.05$?

(e) What conclusions, if any, can be drawn from this study? Answer in ordinary human language.

## Problem 1

A medical school advertises that the mean starting salary of its graduates is $\$89,000$. Concerned that it may actually be less, a group of first-years surveys $38$ recent graduates, finding a sample mean of $\$85,500$. Assume $σ = \$8000$.

**Answer:**

```{r} 
populationMean <- 89000
sampleMean <- 85500
stndDev <- 8000
count <- 38

#PART A
cat("Null-hypotheses:"
    ,"N =",populationMean,"\n")
cat("Alternative-hypotheses:",
    "N <",populationMean,"\n\n")

# PART B
z_score <- (sampleMean - populationMean) / (stndDev / sqrt(count))
cat("z-score:",z_score,"\n\n")

# PART C
p_score <- pnorm(z_score)
cat("p-score:",p_score,"\n\n")

# PART D
if (p_score < 0.05){
  cat("The results are not statistically significant.")
} else if (p_score > 0.05){
  cat("The results are statistically significant.")
} else {
  cat("p-score is equal to 5%.")
}
```

PART E:

That means there is **NOT** enough evidance to support us confidently saying the mean salary of graduates is $89000$ dollars.

## Problem 2

A laptop manufacturer claims that the mean life of the battery for a certain model of laptop is $6$ hours. In a simple random sample of $80$ laptops, the mean battery life is $5.9$ hours. Assume $σ = 1.3$ hours. Is the company’s claim reasonable?

**Answer:**

```{r} 
populationMean <- 6
sampleMean <- 5.9
stndDev <- 1.3
count <- 80

#PART A
cat("Null-hypotheses:"
    ,"N =",populationMean,"\n")
cat("Alternative-hypotheses:",
    "N <",populationMean,"\n\n")

# PART B
z_score <- (sampleMean - populationMean) / (stndDev / sqrt(count))
cat("z-score:",z_score,"\n\n")

# PART C
p_score <- pnorm(z_score)
cat("p-score:",p_score,"\n\n")

# PART D
if (p_score < 0.05){
  cat("The results are not statistically significant.")
} else if (p_score > 0.05){
  cat("The results are statistically significant.")
} else {
  cat("p-score is equal to 5%.")
}
```
PART E:

That means there is enough evidance to support us confidently saying the mean battery life of a phone is $6$ hours.

## Problem 3

A soft drink manufacturer claims that the mean calorie content of one of its sports drinks is $150$ calories per bottle. In a simple random sample of $95$ bottles, the mean is $158$ calories. Is there sufficient evidence to conclude that the mean is actually more than $150$ calories/bottle? Assume $σ = 20$ calories.

**Answer:**

```{r} 
populationMean <- 150
sampleMean <- 158
stndDev <- 20
count <- 95

#PART A
cat("Null-hypotheses:"
    ,"N =",populationMean,"\n")
cat("Alternative-hypotheses:",
    "N >",populationMean,"\n\n")

# PART B
z_score <- (sampleMean - populationMean) / (stndDev / sqrt(count))
cat("z-score:",z_score,"\n\n")

# PART C
p_score <- 1 - pnorm(z_score)
cat("p-score:",p_score,"\n\n")

# PART D
if (p_score < 0.05){
  cat("The results are not statistically significant.")
} else if (p_score > 0.05){
  cat("The results are statistically significant.")
} else {
  cat("p-score is equal to 5%.")
}
```

PART E:

That means there is **NOT** enough evidance to support us confidently saying the mean of calories in a drink is $150$ calories.