---
author: "Sepehr Akbari"
date: "20 march 2024"
format: pdf
---

###### Sepehr Akbari

###### March 20, 2024

# Sampling distributions

## Problem 1

The heights of four friends, in centimeters, are $160$, $165$, $170$, and $185$.

(a) Compute the mean $\mu$ and standard deviation $\sigma$ of their heights.

**Answer:**

```{r}
heights <- c(160,165,170,185)

mean(heights)
```

```{r}
heights <- c(160,165,170,185)

sd(heights)
```

(b) List all possible random samples of size $2$ (with replacement). Compute the sample mean $\overline{x}$ of each. Hint: order doesn’t matter, so there are $10$ possibilities.

**Answer:**

```{r}
heights <- c(160,165,170,185)
num = 1

for (i in 1:length(heights)) {
 for (j in i:length(heights)) {
   samples <- paste0("(", heights[i], ", ", heights[j], ")")
   means <- mean(c(heights[i], heights[j]))
   
   cat("Sample",num,":", samples, ", Mean =",means, "cm\n")
   num = num + 1
 }
}
```

(c) Compute the mean $\mu_{\overline{x}}$ and standard deviation $\sigma_{\overline{x}}$ of these $10$ values of $\overline{x}$. This is the mean and standard deviation of the sampling distribution of the sample mean.

**Answer:**

```{r}
heights <- c(160,165,170,185)
xall <- c()

for (i in 1:length(heights)) {
 for (j in i:length(heights)) {
   samples <- paste0("(", heights[i], ", ", heights[j], ")")
   means <- mean(c(heights[i], heights[j]))
   
   xall <- c(xall,means)
 }
}

mean(xall)
```

```{r}
heights <- c(160,165,170,185)
xall <- c()

for (i in 1:length(heights)) {
 for (j in i:length(heights)) {
   samples <- paste0("(", heights[i], ", ", heights[j], ")")
   means <- mean(c(heights[i], heights[j]))
   
   xall <- c(xall,means)
 }
}

sd(xall)
```

(d) Verify that $\mu_{\overline{x}} =\mu$ and $\sigma_{\overline{x}} = \frac{\sigma}{\sqrt{n}}$ in is case.

**Answer:**

```{r}
heights <- c(160,165,170,185)
xall <- c()

for (i in 1:length(heights)) {
 for (j in i:length(heights)) {
   samples <- paste0("(", heights[i], ", ", heights[j], ")")
   means <- mean(c(heights[i], heights[j]))
   
   xall <- c(xall,means)
 }
}

if(mean(heights) == mean(xall)){
  cat("They are equal")
}else{
  cat("They are NOT equal")
}
```

```{r}
heights <- c(160,165,170,185)
xall <- c()

for (i in 1:length(heights)) {
 for (j in i:length(heights)) {
   samples <- paste0("(", heights[i], ", ", heights[j], ")")
   means <- mean(c(heights[i], heights[j]))
   
   xall <- c(xall,means)
 }
}

size = 2
stndDev = sd(heights)

approximity <- 0.0000000001

if (abs((stndDev / (sqrt(size))) - sd(xall)) < approximity) {
  cat("They are equal")
} else {
  cat("They are NOT equal")
}
```

