---
author: "Sepehr Akbari"
date: "17 April 2024"
format: pdf
---

# Midterm 3: Takehome

## Problem 6

This problem refers to the garbage weight data set, available on Moodle. This set represents a random sample of garbage from $62$ suburban homes.

Loading the dataset (garbage weight):

```{r}
library(readxl)
garbage_weight <- read_excel("/Users/sepehrakbari/Documents/LFC/Semester 2/MATH 150/DSs/garbage_weight.xlsx")
#View(garbage_weight)  
```

(a) Construct a level $99%$ confidence interval for the average amount of glass waste by direct calculation. Identify the point estimate, margin of error, and interval endpoints. Make sure your work is clear.

```{r}
mean <- mean(garbage_weight$GLASS)
standard_deviation <- sd(garbage_weight$GLASS)
cofidence_level <- 0.995
size <- length(garbage_weight$GLASS)
degree_of_freedom <- size - 1

t_star <- qt(cofidence_level, degree_of_freedom)
margin_of_error <- t_star * standard_deviation / sqrt(size)

lower_endpoint <- mean - margin_of_error
upper_endpoint <- mean + margin_of_error

cat("Point Estimate (Mean):",mean,"\n")
cat("Margin of Error:",margin_of_error,"\n")
cat("The Confidence Interval will be from",lower_endpoint,"to",upper_endpoint,"pound. \n")
```

(b) Confirm the results of part (a) with a single line of R code. Include both code and output.

```{r}
t.test(garbage_weight$GLASS, conf.level=0.99)
```

(c) Carefully interpret your answer using the language developed in class.

Based on the data given, about $99\%$ of suburbon households throw away about $2.7$ to $4.8$ pounds of garbage.

## Problem 7

Does the garbage weight set provide evidence that the average food waste per household is more than $3.5$ pounds? Follow all best practices from class and test at significance level $\alpha= 0.05$.

```{r}
cat("Null Hypothesis (H0): H0 = 3.5 \n")
cat("Alternative Hypothesis (H1): H1 > 3.5 (one-sided)\n\n")

population_mean <- mean(garbage_weight$FOOD)
standard_deviation <- sd(garbage_weight$FOOD)
size <- length(garbage_weight$FOOD)
degree_of_freedom <- size - 1 
sample_mean <- 3.5
alpha <- 0.05

t_score <- (population_mean - sample_mean) / (standard_deviation / sqrt(size))
p_value <- 1 - pt(t_score, degree_of_freedom)

cat("t-score is:",t_score,"\n")
cat("p-value is:",p_value,"\n\n")

if(p_value < alpha) {
  cat("There is sufficient evidance to show that
the average food waste per household is greater than 3.5 pounds.")
}else{
  cat("There is insufficient evidance to show that
the average food waste per household is greater than 3.5 pounds.")
}
cat(" We draw no conclusions however.")
```

Or we can use R's t.test() function to calculate or justify our answer:

```{r}
t.test(garbage_weight$FOOD, mu = 3.5, alternative = "greater")
```