---
title: "Simple Linear Regression"
author: "Sepehr Akbari"
date: "11/25/2024"
format: html
embed-resources: true
editor_options: 
  chunk_output_type: console
---

::: {.callout-note appearance="minimal" icon="false"}
The first 5 problems refer to the `modeldata::penguins` data set, which we may view as
representing a simple random sample of antarctic penguins. Adhere to all the best-practices
covered in class.
:::

```{r}
library(tidyverse)
library(modeldata)
```


## Problem 1

::: {.callout-note appearance="minimal" icon="false"}
Create a scatterplot with regression line showing the relationship between bill length and bill
depth for gentoo penguins. Use depth as the explanatory variable. Omit the standard error
ribbon and suppress all output other than the plot and code.
:::

As we are only interested in the gentoo penguins, lets filter our sample to only include them.

```{r}
gentoos <- penguins %>% 
  filter(species == "Gentoo")
```

Now we plot.

```{r}
ggplot(
  gentoos,
  aes(x = bill_depth_mm, 
      y = bill_length_mm)
) +
  geom_point() +
  geom_smooth(method = "lm", 
              se = FALSE
  ) +
  labs(
    title = "Gentoos' Bill Length vs Depth",
    x = "Bill Depth (mm)",
    y = "Bill Length (mm)"
  ) +
  theme_minimal()
```


## Problem 2

::: {.callout-note appearance="minimal" icon="false"}
Determine the equation of this regression line and use it to predict the bill length of a gentoo
penguin with a bill depth of 15.5 millimeters. What is the residual of the observation in this
set with an actual bill depth of 15.5 millimeters?
:::

The equation of a line is in the form of: 

$$ y = mx + b $$

where $m$ is the slope and $b$ is the intercept.

We can get the equation of the regression line by fitting a linear model to the data.

```{r}
model <- lm(bill_length_mm ~ bill_depth_mm, data = gentoos)
summary(model)
```

So the estimated slope is $2.0208$ and the estimated intercept is $17.2295$.

$$ y = 2.0208x + 17.2295 $$




## Problem 3

::: {.callout-note appearance="minimal" icon="false"}
Does the data provide evidence of a linear relationship between `bill_length_mm` and
`bill_depth_mm` among gentoos in the larger population of penguins? Follow all best-practices
from class.
:::

## Problem 4

::: {.callout-note appearance="minimal" icon="false"}
Create an attractive residual plot for the regression model above. Are any trends apparent
that might raise concerns about the model?
::::
























