---
title: "Multiple regression with R"
author: "Sepehr Akbari"
date: "12/6/2024"
format: html
embed-resources: true
editor_options: 
  chunk_output_type: console
---

::: {.callout-note appearance="minimal" icon="false"}
The goal of this set is to model the length of the bills of penguins in the `modeldata::penguins` set.
:::

```{r}
#| message: false
#| warning: false
library(tidyverse)
library(modeldata)
```

## Problem 1

::: {.callout-note appearance="minimal" icon="false"}
Remove all observations that do not include information about the sex of the penguin. Save your result as `penguins_sm` and use it for all questions going forward.
:::

```{r}
#| message: false
#| warning: false
penguins_sm <- penguins %>% 
  filter(!is.na(sex))
```

## Problem 2

::: {.callout-note appearance="minimal" icon="false"}
Create an attractive scatterplot showing bill length on the y-axis (it’s the response variable) and flipper length on the x-axis. Color the points according to sex and add regression lines for the two groups.
:::

```{r}
#| message: false
#| warning: false
ggplot(
  penguins_sm,
  aes(x = flipper_length_mm, 
      y = bill_length_mm,
      color = sex)
) +
  geom_point() +
  geom_smooth(method = "lm", 
              se = FALSE) +
  labs(
    x = "Flipper length (mm)",
    y = "Bill length (mm)",
    title = "Penguins Bill & Flipper Length",
    color = "Sex"
  ) +
  scale_color_brewer(palette = "Dark2") +
  theme_minimal()
```


## Problem 3

::: {.callout-note appearance="minimal" icon="false"}
Build a linear model for bill length as explained by flipper length. What is the equation of the model? What is the adjusted $R^2$ value?
:::

```{r}
#| message: false
#| warning: false
modelBill_flipper <- lm(bill_length_mm ~ flipper_length_mm, 
                         data = penguins_sm)
summary(modelBill_flipper)
```

The equation of the model is:

$$
\text{bill length} = 0.25 (\text{flipper length}) -7.22
$$

The adjusted $R^2$ value is **0.4248**, meaning approximately 42.48% of the variability in bill length is explained by flipper length. The predictor, flipper length, is highly significant (p < 2e-16), indicating a strong relationship with bill length.

## Problem 4

::: {.callout-note appearance="minimal" icon="false"}
Build a linear model for bill length as explained by flipper length and body mass. What is the equation of the model? What is the adjusted $R^2$ value?
:::

```{r}
#| message: false
#| warning: false
modelBill_flipper_mass <- lm(bill_length_mm ~ flipper_length_mm + body_mass_g, 
                              data = penguins_sm)
summary(modelBill_flipper_mass)
```


The equation of the model is:

$$
\text{bill length} = 0.23 (\text{flipper length}) + 0.00055 (\text{body mass}) - 3.98
$$

The adjusted $R^2$ value is **0.4246**, meaning approximately 42.46% of the variability in bill length is explained by flipper length and body mass together. However, the inclusion of body mass does not improve the model significantly (p = 0.342). This suggests that flipper length remains the primary predictor of bill length in this model.

## Problem 5

::: {.callout-note appearance="minimal" icon="false"}
Build a linear model for bill length as explained by body mass and sex. What is the equation of the model? What is the adjusted $R^2$ value?
:::

```{r}
#| message: false
#| warning: false
modelBill_mass_sex <- lm(bill_length_mm ~ body_mass_g + sex, 
                         data = penguins_sm)
summary(modelBill_mass_sex)
```

The equation of the model is:

$$
\text{bill length} = 0.0037 (\text{body mass}) + 1.25 (\text{sex}) + 27.91
$$
$\text{[Where Male = 1, Female = 0]}$

The adjusted $R^2$ value is **0.3542**, meaning approximately 35.42% of the variability in bill length is explained by body mass and sex together. Both predictors are significant:

- Body mass (p < 2e-16) has a small but meaningful positive contribution.

- Sex (p = 0.0197) indicates that males have, on average, a bill length approximately 1.25 mm longer than females, holding body mass constant.

## Problem 6

::: {.callout-note appearance="minimal" icon="false"}
Build a linear model for bill length as explained by flipper length, body mass, and sex. What is the equation of the model? What is the adjusted $R^2$ value?
:::

```{r}
#| message: false
#| warning: false
modelBill_flipper_mass_sex <- lm(bill_length_mm ~ flipper_length_mm + body_mass_g + sex, 
              data = penguins_sm)
summary(modelBill_flipper_mass_sex)
```

The equation of the model is:

$$
\text{bill length} = 0.27  (\text{flipper length}) - 0.00067 (\text{body mass}) + 2.3 (\text{sex}) - 8.08
$$

$\text{[Where Male = 1, Female = 0]}$

The adjusted $R^2$ value is **0.4572**, meaning approximately 45.72% of the variability in bill length is explained by flipper length, body mass, and sex together. However:

- Flipper length (p < 2e-16) is the strongest predictor and significantly contributes to the model.

- Sex (p = 7.27e-6) is also significant, with males having longer bills compared to females.

- Body mass (p = 0.2853) is not significant, suggesting it does not meaningfully contribute to explaining bill length when flipper length and sex are already accounted for.










